---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Loading and preprocessing the data
========================
1) Load the data (i.e. read.csv())
```{r}
activitycsv = unzip("activity.zip")
activity <- read.csv(activitycsv)
```
2) Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
noNa <- na.omit(activity)
```

What is mean total number of steps taken per day?
========================
1) Calculate the total number of steps taken per day
```{r}
tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

2) If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
```{r}
hist(noNa$steps, main = "Histogram of steps per day")
```

3) Calculate and report the mean and median of the total number of steps taken per day

```{r}
summary(activity$steps)["Mean"]
summary(activity$steps)["Median"]
```

What is the average daily activity pattern?
=============================================

1) Make a time series plot (i.e. ```type = "l"```) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
means <- tapply(noNa$steps, noNa$interval, mean)
times <- as.integer(names(means))
plot(times, means, type = 'l', main = "Means vs Time Intervals")
```
2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(sort(tapply(noNa$steps, noNa$interval, mean) ,decreasing = TRUE)[1])
```

Imputing missing values
========================
1) Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
summary(activity$steps)["NA's"]
```


2) Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3)

Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
new <- activity
new$steps[is.na(new$steps)] <- summary(activity$steps)["Mean"]

```
4) Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
hist(new$steps, main = "Histogram of steps per day")
summary(new$steps)["Mean"]
summary(new$steps)["Median"]
```

**Do these values differ from the estimates from the first part of the assignment?** *NO* 

**What is the impact of imputing missing data on the estimates of the total daily number of steps?**  *NO IMPACT*

Are there differences in activity patterns between weekdays and weekends?
============================================
1) Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r results="hide"}
new$w <- weekdays(as.Date(new$date))
for(num in seq_along(new$w)){
    if(new$w[num] == "Saturday" | new$w[num] == "Sunday") new$w[num] <- "weekend"
    else new$w[num] <- "weekday"
}
as.factor(new$w)
weekend <- subset(new, new$w=="weekend")
weekday <- subset(new, new$w=="weekday")
```

```{r}
par(mfrow=c(2,1))
weekendmeans <- tapply(weekend$steps, weekend$interval, mean)
weekendtimes <- as.integer(names(weekendmeans))
weekdaymeans <- tapply(weekday$steps, weekday$interval, mean)
weekdaytimes <- as.integer(names(weekdaymeans))
plot(weekendtimes, weekendmeans, type = 'l', main = "Weekend Means vs Time Intervals")
plot(weekdaytimes, weekdaymeans, type = 'l', main = "Weekday Means vs Time Intervals")
```